Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Introduction.tex
Add TRS intro

Formal verification of cache coherence protocols has traditionally been done
using model checking. That family of analyses significantly restricts the size
as well as complexity of the system that has to be verified; most model-checking
based approaches can not handle more than one level of hierarchy. To the best of
our knowledge, the largest system that has been verified using model checking
has only 12 processors\xxx{the comment was meant to cite other papers from this paper as well. anyways i will take a crack at it}~\cite{Chen:2008:VHC:1559300}, in an era when many
observers are predicting mainstream systems with hundreds of processor cores.

@@

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache-coherent shared memories, starting from a
specification designed to support verification.

>>>>

Formal verification of cache coherence protocols has traditionally been done
using model checking. That family of analyses significantly restricts the size
as well as complexity of the system that has to be verified; most model-checking
based approaches can not handle more than one level of hierarchy. To the best of
our knowledge, the largest system that has been verified using model checking
has only 12 processors~\cite{Chen:2008:VHC:1559300}, in an era when many
observers are predicting mainstream systems with hundreds of processor cores.

@@

Protocols, post verification, are often described via Term Rewrite
Systems (TRS)~\cite{bezem2003term}, which apply equational logic to
describe systems at an interesting point on the spectrum between
declarative and operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after
its verification, can be transliterated into
Bluespec~\cite{Bluespec:TFRG}, a hardware description language based
on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly
implemented into efficient hardware. In this way one reduces the
correctness of the implementation to the correctness of the protocol,
with no need for separate reasoning about an implementation. While the
current Bluespec compiler does not support some implementation
elements that are common in practice (\eg{} fairness guarantees,
atomic actions running over multiple clock cycles), recent work has
provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its
future evolutions represent a very promising platform for productive
implementation of hardware components, including cache-coherent shared
memories, starting from a specification designed to support
verification.
